%% bmc-agent: Agentic Model Checking
%% Conference paper draft — April 2026
%%
%% Build:
%%   pdflatex main && bibtex main && pdflatex main && pdflatex main
%%
%% Target venues: CAV, FSE, ISSTA, ASE

\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit, backgrounds}
\usepackage{hyperref}
\usepackage{abstract}
\usepackage{titlesec}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{authblk}

%% Compact section headings
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalsize\itshape}{\thesubsubsection}{1em}{}

%% ── Tool names ───────────────────────────────────────────────────────────────
\newcommand{\tool}{bmc-agent\xspace}
\newcommand{\fmagent}{FM-Agent\xspace}
\newcommand{\vibeos}{VibeOS\xspace}

%% ── Confidence tiers (inline) ────────────────────────────────────────────────
\newcommand{\dynconf}{\textsc{confirmed\_dynamic}\xspace}
\newcommand{\sysconf}{\textsc{confirmed\_system\_entry}\xspace}
\newcommand{\bmcconf}{\textsc{confirmed\_bmc}\xspace}
\newcommand{\likelyconf}{\textsc{likely}\xspace}

%% ── Listing style ────────────────────────────────────────────────────────────
\lstset{
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\bfseries,
  commentstyle=\itshape\color{gray},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny,
  xleftmargin=1.5em,
  framexleftmargin=1.5em,
}

%% ─────────────────────────────────────────────────────────────────────────────

\title{\textbf{Agentic Model Checking}}

\date{April 2026}

\begin{document}

\maketitle

\begin{abstract}
We present \emph{agentic model checking}, an architecture for automated
verification of systems software that pairs LLM agents with a sound bounded
model checking (BMC) backend. The central principle is \emph{agents propose,
conventional tools dispose}: LLM agents perform specification generation,
counterexample classification, and precondition refinement, but every
soundness-relevant decision they produce is gated by a deterministic check
before affecting the verification verdict.

\tool, our prototype implemented in the AProver project
(\url{https://github.com/theyoucheng/aprover}), operates in four phases: an LLM agent generates
per-function pre/postconditions top-down along the call graph; CBMC checks
each function in isolation against its spec with callee stubs constrained by
their postconditions; a three-way classifier distinguishes confirmed real bugs,
confirmed spurious counterexamples, and unresolved candidates; and a
refinement agent proposes precondition tightenings that pass an SMT soundness
guard before triggering compositional re-verification of all callers.

We evaluate \tool on \vibeos, a bare-metal ARM64 hobby operating system
(\textasciitilde{}15{,}000 lines of LLM-generated C). In raw-source mode with
dynamic validation enabled, \tool reports \textbf{173 confirmed findings}
across 37 kernel modules: 32 faults confirmed at runtime (\dynconf{}),
108 findings with full call chains traced to kernel entry points
(\sysconf{}), and 33 with caller-reachability established by BMC
(\bmcconf{}). An ablation removing the refinement loop
(\emph{FilteringOnly}) reports 59 findings on five wrapper modules vs.\
\tool's 41, showing that refinement acts as a precision filter rather than a
recall enhancer on this corpus.
\end{abstract}

\noindent\textbf{Keywords:} bounded model checking, LLM agents, specification
inference, counterexample classification, compositional verification, systems
software

%% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}
%% ─────────────────────────────────────────────────────────────────────────────

LLM coding assistants now generate substantial bodies of systems-level code,
including operating systems, device drivers, and embedded utilities. Deployed
artefacts from these tools raise a verification challenge that existing methods
address poorly. Developers who deploy LLM-generated code frequently do not
understand every function in detail; writing manual specifications is infeasible
at scale; and the bugs that arise tend to be the low-level kind---integer
overflow, off-by-one indexing, missing bounds checks on malformed input---that
require exhaustive path analysis to verify: an LLM can flag suspicious
patterns but cannot prove their absence across all inputs.

Two recent lines of work bracket the problem.
\fmagent~\cite{fmagent} generates Hoare-style pre/postconditions top-down
from caller context and uses an LLM to reason about each function against
its specification, confirming bugs by running live test cases. It scales
to large systems---applying to a 15k-LoC, 452-function operating
system---and is effective at semantic and logic bugs. Because verification
is performed by an LLM rather than a solver, it is not sound for low-level
arithmetic and memory-safety properties: bugs that require exhaustive
reasoning over all input ranges can escape undetected.
AutoSpec~\cite{autospec} and Lemur~\cite{lemur} couple LLM-generated
specifications with formal checkers but depend on deductive infrastructure
that does not produce concrete counterexamples and struggles with
loop-heavy systems code.

\tool occupies a different point in this space. Like \fmagent, it generates
specifications top-down so each function's expected behavior is derived from
its callers. Like AutoSpec and Lemur, it uses a sound solver for the checking
step. Unlike all three, it uses \emph{bounded model checking} as the solver,
which brings three structural advantages:

\begin{enumerate}
\item \textbf{Concrete counterexamples.} When BMC finds a property violation it
  produces a concrete execution witness, enabling a classifier that
  distinguishes real bugs from artifacts of the abstraction by running the
  witness against the real system.

\item \textbf{Per-function parallelism.} Each function is checked in isolation
  against its spec with callees replaced by spec-constrained stubs, avoiding
  whole-program state-space explosion.

\item \textbf{Meaningful compositional refinement.} When a spurious
  counterexample triggers precondition tightening, the refined spec immediately
  improves verification of every caller---because the refined spec is re-used
  as the callee's stub in every caller's harness.
\end{enumerate}

The architectural principle unifying these choices is \emph{agents propose,
conventional tools dispose}: every LLM output that could affect a verification
verdict passes through a deterministic check (a DSL parser, a CBMC reachability
query, an SMT soundness guard, or concrete execution) before being allowed to
change the verification state. This principle is not new in spirit, but
\tool's instantiation of it---a specific assignment of tasks to agents versus
solvers, connected by a compositional refinement loop---is, to our knowledge,
novel.

\smallskip\noindent\textbf{Contributions.}
\begin{enumerate}
\item The \emph{agentic model checking} architecture: a principled decomposition
  of the verification workflow into agentic and conventional components,
  connected by a compositional CEGAR-style refinement loop
  (\S\ref{sec:architecture}).

\item A \emph{three-way counterexample classifier} with four evidence tiers
  (\dynconf, \sysconf, \bmcconf, \likelyconf),
  supporting nuanced reporting without silently suppressing findings
  (\S\ref{sec:classifier}).

\item A \emph{compositional refinement loop} with an SMT soundness guard that
  prevents over-refinement from suppressing real bugs, and propagates spec
  improvements to callers (\S\ref{sec:refinement}).

\item A prototype implementation targeting C/CBMC with bare-metal kernel
  support, and an evaluation on \vibeos showing 173 confirmed findings across
  37 modules with 32 confirmed at runtime (\S\ref{sec:evaluation}).
\end{enumerate}

%% ─────────────────────────────────────────────────────────────────────────────
\section{Background}
\label{sec:background}
%% ─────────────────────────────────────────────────────────────────────────────

\subsection{Bounded Model Checking}

Bounded model checking (BMC)~\cite{bmc-survey} checks whether a program
property is violated within an execution of at most $k$ steps ($k$ being the
\emph{unwinding bound}). CBMC~\cite{cbmc} applies BMC to C programs: given a
harness that constrains program inputs and asserts output properties, CBMC
unrolls loops up to bound $k$, encodes the resulting path set as a SAT/SMT
formula, and either verifies the property or returns a concrete
counterexample. The soundness guarantee is conditional: \emph{no violation in
executions of depth at most $k$}, not full correctness.

\subsection{Compositional Verification and Assume-Guarantee}

Compositional verification~\cite{assume-guarantee} decomposes a system into
components and verifies each in isolation, with inter-component interactions
captured by \emph{contracts}---assume-guarantee pairs. A component is
verified under the assumption that its environment satisfies its input
contract; the proof obligation is that the component produces outputs
satisfying its output contract. When contracts are sound (they correctly
characterize actual behavior), individual component proofs compose into a
whole-system proof. \tool instantiates this framework: each function's
pre/postconditions are the contract, and BMC checks each function against its
contract with callee stubs constrained by the callees' postconditions.

Counterexample-guided abstraction refinement
(CEGAR)~\cite{cegar} iteratively refines an abstract model when a spurious
counterexample is found. \tool applies the same idea at the specification
level: a spurious counterexample triggers a precondition tightening that
removes the spurious behavior, with CBMC re-verifying the function under the
refined spec.

\subsection{LLM Agents for Verification}

LLM agents augment language models with tool use---the ability to invoke
external programs and observe their outputs within a reasoning
loop~\cite{react}. Recent work has explored LLM agents for verification tasks
including specification generation~\cite{fmagent,autospec,lemur} and
loop invariant inference~\cite{wu-invariants}. \tool contributes to this line by combining
an agentic layer with a sound solver in a closed compositional refinement loop.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Architecture}
\label{sec:architecture}
%% ─────────────────────────────────────────────────────────────────────────────

\subsection{Overview}

Figure~\ref{fig:architecture} presents the \tool pipeline. Two phases are
agentic (shaded): the \emph{spec generator} and the \emph{CEx validator +
refiner}. Two are conventional: the \emph{BMC engine} and the \emph{soundness
guard}. The dashed arc is the refinement feedback loop.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance = 5mm and 8mm,
  box/.style    = {draw, rounded corners=3pt, minimum width=2.4cm,
                   minimum height=0.7cm, align=center, font=\small},
  agent/.style  = {box, fill=violet!15, draw=violet!60},
  conv/.style   = {box, fill=gray!12,   draw=gray!60},
  arr/.style    = {-{Stealth[length=4pt]}, thick},
  darr/.style   = {arr, dashed},
]

%% Column 1
\node[agent] (sg)   {Spec Generator\\{\scriptsize (LLM)}};

%% Column 2
\node[conv,  right=of sg]  (sched)  {Scheduler\\{\scriptsize (call graph)}};

%% Column 3
\node[conv,  right=of sched] (bmc) {BMC Engine\\{\scriptsize (CBMC)}};

%% Column 4 — stacked
\node[agent, right=of bmc]        (cls) {CEx Classifier\\{\scriptsize (LLM + BMC)}};
\node[agent, below=of cls]        (ref) {Refiner\\{\scriptsize (LLM)}};
\node[conv,  below=of ref]        (sg2) {Soundness Guard\\{\scriptsize (SMT)}};

%% Outputs
\node[right=7mm of cls, align=left, font=\small]
  (bugs) {Bug\\Reports};

%% Arrows
\draw[arr] (sg)    -- (sched);
\draw[arr] (sched) -- (bmc);
\draw[arr] (bmc)   -- node[above,font=\scriptsize]{CEx} (cls);
\draw[arr] (cls)   -- node[right,font=\scriptsize]{spurious} (ref);
\draw[arr] (ref)   -- (sg2);
\draw[arr] (cls)   -- node[above,font=\scriptsize]{real} (bugs);

%% Feedback arc
\draw[darr] (sg2.west)
  .. controls +(-1.6,0) and +(-0.5,-1.4) ..
  node[left,font=\scriptsize,align=right]{refined\\spec}
  (sched.south);

%% Source input
\node[left=6mm of sg, font=\small, align=right] (src) {C source\\+ context};
\draw[arr] (src) -- (sg);

\end{tikzpicture}
\caption{\tool pipeline. Violet (shaded) boxes are LLM-driven; gray boxes are
deterministic. The dashed arc is the refinement feedback loop: when a
counterexample is confirmed spurious, the refiner proposes a precondition
tightening that passes an SMT soundness guard and re-triggers verification of
the refined function and all its callers.}
\label{fig:architecture}
\end{figure}

\subsection{Phase 1: Specification Generation}
\label{sec:specgen}

The spec generator is an LLM agent that traverses the call graph of the input
C source top-down, starting from functions that have no callers (entry points
or top-layer functions), and working layer by layer toward leaf functions.
Generating specs top-down means each function's expected behavior is derived
from how its callers invoke it---capturing caller intent---rather than from the
potentially-buggy implementation itself.

For each function, the agent:
\begin{enumerate*}[label=(\roman*)]
  \item retrieves the function body, its callers' bodies, and any available
  domain knowledge (API documentation, datasheet excerpts);
  \item drafts a pre/postcondition pair in a restricted DSL;
  \item validates the draft against a deterministic parser; and
  \item retries on failure, up to a configurable limit.
\end{enumerate*}

\smallskip\noindent\textbf{Dual-source generation.}
When enabled, \tool generates each spec twice with different prompts---one
emphasizing the function's role from the callers' perspective, one emphasizing
its implementation behavior---and flags disagreements for human review.
Disagreement is a useful signal that the function's intended semantics is
ambiguous or that the LLM is uncertain.

\smallskip\noindent\textbf{Specification DSL.}
Specifications are expressed in a small DSL designed to be simultaneously
expressive enough for driver- and OS-style bugs, reliably emittable by an LLM,
and directly translatable to CBMC constructs without interpreter intervention.
Core predicates include \lstinline{valid(ptr)} (non-null, allocated),
\lstinline{valid_string(ptr)} (null-terminated), and
\lstinline{valid_range(ptr, lo, hi)} (in-bounds range). Return-value
constraints are first-order formulas over integers and pointer fields. The DSL
excludes uninterpreted functions, which eliminates a common LLM failure
mode---generating specifications that refer to concepts the solver cannot
evaluate.

The DSL translator converts each spec into \lstinline{__CPROVER_assume} and
\lstinline{__CPROVER_assert} calls within the synthesized harness. This
translation is a deterministic template substitution, not an LLM call.

\smallskip\noindent\textbf{Strongly connected components.}
Functions involved in mutual recursion are condensed into a single SCC node
and specified and verified as a unit rather than individually. The SCC
condensation is computed by Kosaraju's algorithm on the call graph.

\subsection{Phase 2: BMC Engine}
\label{sec:bmc}

For each function $F$, the BMC engine synthesizes a CBMC harness that:
\begin{enumerate*}[label=(\roman*)]
  \item nondeterministically initializes $F$'s parameters subject to its
    precondition via \lstinline{__CPROVER_assume};
  \item provides a stub for each callee $G$ of $F$ that nondeterministically
    returns a value satisfying $G$'s postcondition, also via
    \lstinline{__CPROVER_assume};
  \item asserts $F$'s postcondition at all return sites; and
  \item injects CBMC's built-in safety properties (null dereference, integer
    overflow, out-of-bounds array access, division by zero).
\end{enumerate*}
CBMC is invoked under a configured loop unwinding bound $k$ (default 4) and
per-function timeout (default 120 s). Verdicts are either \emph{verified}
(no violation within bound $k$) or a concrete counterexample with full
variable assignments.

Functions are dispatched to CBMC in call-graph order---leaf functions first,
callers after all their callees have been assigned a spec. This ensures that
each callee's spec is available when a caller's harness is synthesized.
Functions within the same call-graph layer are dispatched in parallel.

\subsection{Phase 3: Counterexample Classification}
\label{sec:classifier}

A central design decision in \tool is that a CBMC counterexample is
\emph{not} treated as a bug report directly. Every counterexample passes
through a three-way classifier that produces one of:

\begin{description}
\item[\textbf{Real bug.}] Reachability and feasibility checks confirm the
  violation is reachable under the real program; reported with one of four
  evidence tiers (Table~\ref{tab:tiers}).
\item[\textbf{Spurious.}] No caller can supply inputs that produce the
  violating state; the counterexample is an artifact of stub
  over-approximation. Passed to the refiner.
\item[\textbf{Unresolved.}] Checks are inconclusive in both directions.
  Reported as a candidate for human review; never silently dropped.
\end{description}

The three-way structure is important for soundness: a two-way classifier must
commit on every counterexample, and misclassifying a real bug as spurious
suppresses it silently. \emph{Unresolved} is the honest ``cannot confirm''
verdict.

\subsubsection{Stage 1: Input reachability}

For a counterexample at function $F$ with witness input $\mathbf{x}$, Stage 1
asks: \emph{can any caller of $F$ actually supply $\mathbf{x}$?}
A CBMC harness runs each in-file caller $C$ with its other callees stubbed
(constrained by their postconditions) and $F$'s argument fixed to
$\mathbf{x}$; CBMC checks whether $C$ has an execution that produces
$\mathbf{x}$ at the call site. If no in-file callers exist but cross-file
callers are known (whole-codebase mode), the same harness is run against
those callers in their own file context. If $F$ has no callers anywhere it is
a system entry point and Stage 1 is skipped entirely.

After at least one caller confirms reachability, \tool runs
\emph{upward propagation}: it recurses through the confirming caller's own
callers, then their callers, and so on, trying to reach a function with no
callers anywhere in the analyzed corpus (a true system entry point). Whether
this propagation reaches the top determines the final tier.

\subsubsection{Stage 2: Callee feasibility}

Stage 1 confirms that $F$'s inputs are caller-reachable, but the callee
return values $R_1, R_2, \ldots$ in the witness trace may be unrealistic:
a stub can return any value satisfying its postcondition, including values
that no real implementation ever produces. Stage 2 checks this by inlining
the real bodies of $F$'s locally-available callees, fixing $F$'s scalar
inputs to the witness values, and re-running CBMC. If the violation still
fires with real callee bodies it is feasible; if CBMC verifies instead, the
violation cannot occur with real callees and the counterexample is
reclassified as \emph{unresolved}.

Stage 2 is applied in two situations: (i) $F$ is a system entry point (no
callers anywhere, Stage 1 skipped), and (ii) Stage 1 confirmed reachability
from at least one \emph{in-file} caller. It is \emph{not} applied when
reachability was confirmed only via a cross-file caller.

\subsubsection{Stage 3: Dynamic validation}

After any real-bug verdict, \tool optionally compiles a GCC harness that
uses the highest confirmed caller in the call chain as the entry point, with
the witness input values hardcoded and signal handlers catching SIGSEGV,
SIGABRT, SIGFPE, and SIGILL. Executing the harness on the host machine
provides ground-truth confirmation independent of all model assumptions; a
caught signal upgrades the finding to \dynconf regardless of its current
tier.

Bare-metal sources require compatibility transformations before host
compilation: stripping ARM64 inline assembly, removing \texttt{static inline}
libc stubs that conflict with system headers, removing glibc-internal typedef
redefinitions, and stripping non-standard function re-declarations. A
fallback link with \verb|-Wl,--unresolved-symbols=ignore-all| handles
functions referencing globals from other translation units.

\subsubsection{Tier assignment}

Table~\ref{tab:tiers} summarises how the stage outcomes map to evidence
tiers. The tier assignment rule (highest matching tier wins) is:
\begin{enumerate}
  \item Stage 3 caught a signal $\Rightarrow$ \dynconf.
  \item Upward propagation reached a no-caller entry point
        $\Rightarrow$ \sysconf.
  \item Stage 1 (in-file) + Stage 2 passed, propagation stopped mid-chain
        $\Rightarrow$ \bmcconf.
  \item Over-refinement guard triggered $\Rightarrow$ \likelyconf.
\end{enumerate}

\begin{table}[t]
\caption{Evidence tiers for confirmed real bugs.}
\label{tab:tiers}
\small
\begin{tabular}{p{2.6cm}p{4.9cm}}
\toprule
\textbf{Tier} & \textbf{Checks performed} \\
\midrule
\dynconf
  & Stage 3: GCC harness with hardcoded witness executed on host; signal
    (SIGSEGV/SIGABRT/SIGFPE/SIGILL) caught. No model assumptions. \\[4pt]
\sysconf
  & Full call chain traced to a no-caller entry point.
    \emph{Sub-case A} (entry function): Stage 2 callee inlining run and
    passed; violation fires with real callees.
    \emph{Sub-case B} (cross-file reachability): Stage 1 cross-file CBMC
    harness confirms reachability; Stage 2 not run. \\[4pt]
\bmcconf
  & Stage 1 in-file CBMC reachability passed (at least one in-file caller
    confirms) + Stage 2 callee inlining passed; upward propagation did not
    reach a no-caller entry point. \\[4pt]
\likelyconf
  & No Stage 1/2 confirmation; over-refinement guard triggered, indicating
    the violation is likely real but not formally confirmed. \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Phase 3 (cont.): Refinement Loop}
\label{sec:refinement}

When a counterexample is confirmed spurious, the \emph{refiner}---an LLM
agent---performs blame analysis to identify which component's specification is
responsible, and proposes a tightened precondition for function $F$.

\smallskip\noindent\textbf{Blame analysis.}
The refiner runs $F$'s real implementation on the counterexample's concrete
input, compares the observed behavior against each spec in the harness, and
identifies the inconsistency: Was $F$'s precondition too weak (allowing
inputs that violate the spec's intent)? Was a callee's postcondition
over-permissive (claiming the callee can return values it never produces)?
Was $F$'s postcondition too strong (requiring more than $F$ actually
guarantees)? The blame identifies where the spec needs tightening.

\smallskip\noindent\textbf{Soundness guard.}
A proposed precondition tightening is not applied directly. The soundness
guard first checks via SMT whether the tightening would exclude any state
actually reachable from $F$'s callers. If it would, the refinement is
rejected and the counterexample is reported as \emph{unresolved} rather than
refined away. This prevents the refinement loop from silently suppressing
real bugs that happen to share the input space with spurious ones.

\smallskip\noindent\textbf{Compositional propagation.}
Once a refinement is accepted, the scheduler re-queues not only $F$ but every
function that calls $F$. This reflects the compositional structure: $F$'s
spec is used as $F$'s stub in every caller's harness. When the spec changes,
each caller must be re-verified against the new stub. This propagation is the
primary justification for the refinement loop: local filtering (discarding
the spurious counterexample without updating the spec) would suppress the
false alarm in $F$ but would not improve verification of $F$'s callers.

Refinement can cascade: a new counterexample discovered during a caller's
re-verification may itself be spurious, triggering further refinement up the
call graph. A per-counterexample iteration cap (default 5) and a
per-function re-queue limit bound the total work.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Soundness}
\label{sec:soundness}
%% ─────────────────────────────────────────────────────────────────────────────

\tool is \emph{sound per function within the BMC unwinding bound, conditional
on spec correctness and on the faithfulness of any LLM-synthesized environment
stubs}.

Formally: if CBMC returns \textsc{verified} for function $F$ under bound $k$,
then for every execution of $F$ starting from a state satisfying $F$'s
precondition, in which every callee of $F$ behaves according to its
specification, and in which no loop iterates more than $k$ times, $F$
satisfies its postcondition and all injected safety properties.

The conditionality matters: an incorrect LLM-generated spec---either
over-strong (causing spurious counterexamples) or over-weak (causing
silently-missed bugs)---causes CBMC to verify the function against the wrong
contract. Over-strong specs are handled by the classifier and refiner.
Over-weak specs represent the primary remaining soundness gap; the
spec-quality defenses (\S\ref{sec:impl}) are the intended mitigation.

The \emph{agents propose, conventional tools dispose} principle is directly
auditable:
(i) LLM-generated specs are validated by a deterministic DSL parser before
reaching CBMC;
(ii) LLM-proposed refinements are gated by the SMT soundness guard;
(iii) LLM classifications are cross-checked by a CBMC reachability query.
Where LLM output is not gated by a conventional check---most critically, the
initial spec generation---the soundness claim weakens correspondingly.

\smallskip\noindent\textbf{Asymmetry in Phase 3.}
The \textsc{spurious} direction is sound under the stub model: if even
maximally permissive (fully nondeterministic) callee stubs cannot produce the
counterexample state at any caller, the real, more-constrained callees also
cannot. The \textsc{real\_bug} direction is conditionally sound: input
reachability is a necessary but not sufficient condition. The missing piece
is joint feasibility of the callee return values in the trace---whether all
callees can simultaneously produce the values in the witness. Stage 2 (local callee inlining) addresses this for findings classified via
in-file callers or entry-point paths; it is not applied on the cross-file
reachability sub-path of \sysconf{}. Stage 3 (dynamic execution) provides
ground-truth confirmation for the subset that compiles and runs.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Implementation}
\label{sec:impl}
%% ─────────────────────────────────────────────────────────────────────────────

\tool is implemented in the AProver project, available at
\url{https://github.com/theyoucheng/aprover}. It is built in Python 3.11
using Claude~\cite{anthropic-claude} (claude-sonnet-4-6) as the LLM backend. The core pipeline is 4{,}000 lines; the harness generator,
DSL translator, and dynamic validator add another 4{,}000 lines. The test
suite has 211 passing tests. Table~\ref{tab:components}
summarizes the implementation status of each architectural component.

\begin{table}[t]
\caption{Component implementation status.}
\label{tab:components}
\small
\begin{tabular}{ll}
\toprule
\textbf{Component} & \textbf{Status} \\
\midrule
Spec Generator (LLM)              & Implemented \\
BMC Engine (CBMC)                 & Implemented \\
CEx Classifier: input reachability (Stage 1) & Implemented \\
CEx Classifier: callee inlining (Stage 2)    & Implemented \\
CEx Classifier: dynamic validation (Stage 3) & Implemented \\
Refiner + SMT soundness guard     & Implemented \\
Compositional caller propagation  & Implemented \\
Cross-file call-graph construction& Implemented \\
Spec-quality defenses             & Implemented (not yet run at scale) \\
Kani (Rust) backend               & Scaffolded only \\
\bottomrule
\end{tabular}
\end{table}

\smallskip\noindent\textbf{Spec-quality defenses.}
\tool implements four spec-quality mechanisms that can be enabled at analysis
time: (i) a \emph{coverage checker} that verifies the spec references the
return value, mutated fields, and key parameters; (ii) a \emph{mutation
tester} that injects off-by-one, comparison-flip, and null-check-drop
mutations into the function body and checks via BMC whether the spec catches
the mutant; (iii) a \emph{consistency checker} (LLM-based) that compares
caller-callee spec pairs for logical consistency; and (iv) an \emph{executable
sanity checker} (LLM-based) that simulates concrete inputs against the spec.
These defenses are the intended mitigation for over-weak LLM-generated specs;
running them at scale on the evaluation corpus is future work.

\smallskip\noindent\textbf{Whole-codebase mode.}
The \verb|verify-dir| command preprocesses every \texttt{.c} file in a
directory using the C preprocessor, runs a two-pass global call-graph
construction to identify cross-file caller/callee relationships, and verifies
each file. Functions whose callers reside in other files are not
misclassified as system entry points; instead, \tool runs CBMC reachability
harnesses against those cross-file callers and, if the chain reaches a true
entry function (no callers anywhere in the corpus), promotes the finding to
\sysconf.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Evaluation}
\label{sec:evaluation}
%% ─────────────────────────────────────────────────────────────────────────────

We evaluate \tool on \vibeos~\cite{vibeos}, a bare-metal ARM64 hobby
operating system of approximately 15{,}000 lines of C and 452 functions,
written with substantial LLM assistance. We chose \vibeos because: (i) it is representative of the
LLM-generated systems code that motivates the work; (ii) it is publicly
available and independently developed, so our evaluation is not circular;
and (iii) its bare-metal character stresses harness generation and
dynamic validation in ways that a user-space target would not.

We ask three research questions:

\begin{description}
\item[\textbf{RQ1}] Does \tool find real bugs in real-world C?
\item[\textbf{RQ2}] What is the strength of evidence for reported findings?
\item[\textbf{RQ3}] Does refinement add value over simple filtering of
  spurious counterexamples?
\end{description}

\noindent\textbf{Setup.} CBMC 5.95.1, loop unwinding bound $k=4$, per-function
timeout 120\,s, Claude claude-sonnet-4-6. Dynamic validation uses GCC 12.3 on
an x86-64 Linux host. All runs use whole-codebase mode (\verb|verify-dir|)
except the ablation, which uses per-module wrapper files.

\subsection{RQ1: Synthetic Targets}

To validate the pipeline end-to-end with known ground truth, we first apply
\tool to four synthetic C programs each containing a planted bug:

\begin{center}
\small
\begin{tabular}{llr}
\toprule
Source & Bug & Findings \\
\midrule
\texttt{simple\_driver.c} & off-by-one in ring-buffer write & 3 \\
\texttt{sensor\_hub.c}    & integer overflow in reading accumulation & 2 \\
\texttt{block\_device.c}  & integer overflow in block seek & 1 \\
\texttt{memory\_allocator.c} & null dereference on free & 1 \\
\bottomrule
\end{tabular}
\end{center}

All planted bugs are correctly found and none of the reported findings are
false positives in the sense of not corresponding to any actual reachable
violation. This is a pipeline sanity check, not a claim of precision on
uncontrolled targets.

\subsection{RQ1 + RQ2: VibeOS Full Kernel}

Table~\ref{tab:vibeos-full} shows results from the raw-source mode run with
dynamic validation enabled across all 38 unique kernel module stems (47 of 48
\texttt{.c} files; \texttt{font.c} was not processed in any run). \tool
reports 173 confirmed findings across 37 modules in three evidence tiers.

\begin{table}[t]
\caption{VibeOS raw-source findings by evidence tier.}
\label{tab:vibeos-full}
\begin{tabular}{lr}
\toprule
\textbf{Evidence tier} & \textbf{Count} \\
\midrule
\dynconf{}     & 32  \\
\sysconf{}     & 108 \\
\bmcconf{}     & 33  \\
\midrule
\textbf{Total} & \textbf{173} \\
\bottomrule
\end{tabular}
\end{table}

The 32 \dynconf findings each triggered a SIGSEGV in a GCC-compiled harness
running the exact counterexample witness, confirming the fault independently
of any CBMC model. They span 14 modules including the IRQ handler, USB stack,
ELF loader, VFS, DMA layer, cursor renderer, and TTF font renderer.

\begin{table}[t]
\caption{\dynconf findings by module.}
\label{tab:dynamic-findings}
\small
\begin{tabular}{lp{5.5cm}}
\toprule
\textbf{Module} & \textbf{Functions} \\
\midrule
\texttt{irq}      & \texttt{handle\_serror}, \texttt{handle\_sync\_exception},
                    \texttt{uart\_puthex}, \texttt{wsod\_animate\_ekg},
                    \texttt{wsod\_draw\_line}, \texttt{wsod\_draw\_sad\_mac},
                    \texttt{wsod\_draw\_text}, \texttt{wsod\_hex},
                    \texttt{wsod\_int} \\
\texttt{ttf}      & \texttt{stbtt\_\_sort\_edges\_ins\_sort},
                    \texttt{stbtt\_\_sort\_edges\_quicksort} \\
\texttt{dma}      & \texttt{hal\_dma\_copy}, \texttt{hal\_dma\_fb\_copy} \\
\texttt{fb}       & \texttt{find\_ramfb\_selector}, \texttt{fw\_cfg\_read},
                    \texttt{local\_strcmp} \\
\texttt{usb\_hid} & \texttt{kbd\_ring\_push}, \texttt{mouse\_ring\_push},
                    \texttt{usb\_irq\_handler} \\
\texttt{vfs}      & \texttt{readdir\_callback}, \texttt{vfs\_lookup} \\
\texttt{elf}      & \texttt{elf\_load}, \texttt{elf\_load\_at} \\
\texttt{cursor}   & \texttt{draw\_cursor\_at}, \texttt{save\_background} \\
Others            & \texttt{scroll\_up} (console), \texttt{strlen} (string),
                    \texttt{uart\_getc\_blocking} (kernel),
                    \texttt{process\_init}, \texttt{usb\_enumerate\_hub},
                    \texttt{kapi\_print\_hex}, \texttt{kapi\_print\_int} \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Bug categories.}
Findings fall into two categories. \emph{Unbounded loops}: CBMC's unwinding
assertion (\texttt{unwind.0}) fires when a loop takes \emph{more than $k=4$
iterations} for some concrete input---it does not, by itself, mean the loop is
truly infinite.  Not all \texttt{unwind.0} firings are bugs: some loops
terminate correctly but legitimately require more than four iterations (e.g.,
\texttt{usb\_enumerate\_hub} iterating once per USB port needs five iterations
for a five-port hub).  \tool filters such conservative-bound artifacts
automatically by re-running the Phase~2 harness at a higher bound $k'=16$ with
nondeterministic inputs: if CBMC verifies the function under all valid inputs at
$k'$, the finding is discarded as a conservative-bound artifact; if the
unwinding assertion persists, the loop is genuinely unbounded.

Findings that survive the escalation check represent real bugs.
Two VibeOS examples: passing a non-null-terminated buffer
(e.g., \texttt{char buf[4]=\{'a','b','c','d'\}}) to any string function such
as \texttt{strlen} causes the function to scan past the array end---yielding
an out-of-bounds read and, in the dynamic harness, a
\texttt{SIGSEGV}---because the null terminator is never found.
In \texttt{dtb\_parse}, a malformed device-tree string field whose stored
length is \texttt{UINT\_MAX} causes approximately $2^{32}$ iterations,
effectively hanging the kernel.  Both cases persist at $k'=16$; their
practical severity depends on whether callers can supply the triggering
input---a relevant concern for \texttt{dtb\_parse}, which consumes
externally-supplied firmware data.

\emph{Memory-safety bugs}: the clearest example is signed integer overflow in
\texttt{print\_signed}, where the expression \texttt{num = -num} is undefined
behavior when \texttt{num == INT64\_MIN}. Any call to
\texttt{printf("\%ld", (long)INT64\_MIN)} triggers this UB unconditionally;
the finding is unambiguous.

\noindent\textbf{Cross-validation.} At the time of evaluation, the \vibeos
public issue tracker contained a filed defect (issue~\#26) describing a
\texttt{calloc} integer-overflow vulnerability in the memory allocator.
\tool independently reported the same defect in
\texttt{vibeos\_memory.c} via BMC over a per-function harness. Independent
discovery of the same bug by a different method is a meaningful positive
signal that the classifier is calibrated correctly on at least this finding.

\noindent\textbf{Comparison with FM-Agent.}
\fmagent~\cite{fmagent} also evaluates on \vibeos and reports 23 bugs,
all confirmed by running generated test cases against the live system
(up to 10 attempts per candidate). \tool reports 173 findings, but the
two numbers are \emph{not directly comparable}.

The tools target overlapping but distinct bug classes.
\fmagent finds semantic and logic bugs---incorrect return values, wrong
conditional logic, behavioral errors---confirmed by running test cases
against the live system. \tool targets low-level C undefined
behaviour---null dereference, integer overflow, unbounded
loops---confirmed by BMC and optionally by GCC-harness execution.
\fmagent's 23 bugs include 4 memory-corruption cases, 2 infinite loops,
4 incorrect return values, and 12 logic errors. \tool's findings include
loop-termination violations (e.g., passing a non-null-terminated buffer to
a string function, or a malformed \texttt{UINT\_MAX} length field to
\texttt{dtb\_parse}) and memory-safety violations; these are real bugs,
but their practical severity depends on whether callers can supply the
triggering inputs.

The most comparable subset is \tool's 32 \dynconf findings (runtime
SIGSEGV in a GCC-compiled harness) against \fmagent's 23 runtime-confirmed
bugs; even here the confirmation mechanisms differ---\tool compiles an
x86-64 harness around an isolated function, while \fmagent runs
system-entry test cases against the full kernel. A meaningful head-to-head
comparison would require auditing both result sets manually on the same
modules. The tools are best understood as \emph{complementary}: \fmagent
excels at semantic bugs observable through system execution; \tool excels
at low-level memory-safety violations detectable by bounded model checking.

\subsection{RQ3: Refinement vs.\ Filtering}

To assess the value of the refinement loop (RQ3), we compare \tool in full
mode against a \emph{FilteringOnly} ablation that classifies counterexamples
as real or spurious and drops the confirmed-spurious ones, but never updates
any spec and never re-queues any caller. We run both on five hand-prepared
\vibeos wrapper modules.

\begin{table}[t]
\caption{Full \tool vs.\ FilteringOnly on five \vibeos modules.}
\label{tab:ablation}
\small
\begin{tabular}{lrrr}
\toprule
\textbf{Module} & \textbf{\tool} & \textbf{FilteringOnly} & \textbf{$\Delta$} \\
\midrule
\texttt{vibeos\_vfs.c}    & 18 & 27 & $+9$ \\
\texttt{vibeos\_string.c} & 16 & 22 & $+6$ \\
\texttt{vibeos\_dtb.c}    &  4 &  4 & $\phantom{+}0$ \\
\texttt{vibeos\_printf.c} &  3 &  6 & $+3$ \\
\texttt{vibeos\_klog.c}   &  0 &  0 & $\phantom{+}0$ \\
\midrule
\textbf{Total}            & \textbf{41} & \textbf{59} & $+18$ \\
\bottomrule
\end{tabular}
\end{table}

FilteringOnly reports 44\% more findings (59 vs.\ 41). The direction is
noteworthy: one might expect refinement to \emph{add} bugs by propagating
tightened preconditions to callers and uncovering new violations. On this
corpus, refinement acts primarily as a \emph{precision filter}---it eliminates
18 findings that FilteringOnly reports as real, without adding any unique
findings via caller re-queuing. Two modules (\texttt{dtb}, \texttt{klog}) are
unaffected, suggesting the filtering effect concentrates in the more heavily
interconnected modules.

The mechanistic explanation is that refinement tightens a precondition in
response to one spurious counterexample, which then also excludes the inputs
that trigger other violations in the same function---violations that
FilteringOnly correctly identifies as reachable under the original, weaker
spec. Whether those 18 eliminated findings are genuine precision improvements
(the refined spec correctly rules out inputs that callers never produce) or
over-refinement artifacts (real bugs suppressed by an overly tight spec)
remains the central open question. Resolving it requires a manual audit of
the eliminated findings---the most important next step in the evaluation.

\subsection{Threats to Validity}

\noindent\textbf{Classifier precision.}
The 173 finding count is \tool's own classification, not an externally-audited
ground truth. A manual audit of a randomly-sampled subset is needed to
establish the false-positive rate of the \textsc{real\_bug} verdict.

\noindent\textbf{Unwinding bound.}
The default bound $k=4$ is conservative and favors loop-termination findings
over deeper intra-loop bugs. The single unambiguous memory-safety finding
(\texttt{print\_signed}) is not sensitive to $k$; the 40 unbounded-loop
findings are.

\noindent\textbf{Single-system evaluation.}
All real-world results are from a single target (\vibeos). Generalization to
other LLM-generated codebases, historical buggy driver corpora, or
non-ARM-targeted kernels is not yet demonstrated.

\noindent\textbf{Spec correctness.}
Soundness is conditional on LLM-generated specs being correct. Over-weak
specs---the primary correctness failure mode---cause bugs to be missed
silently. The spec-quality defenses (\S\ref{sec:impl}) are the intended
mitigation and have not yet been run at scale.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Related Work}
\label{sec:related}
%% ─────────────────────────────────────────────────────────────────────────────

\noindent\textbf{LLM agents for verification.}
\fmagent~\cite{fmagent} is the closest prior work: it generates Hoare-style
pre/postconditions top-down from caller context and uses LLM reasoning to
verify each function against its spec, scaling to the same \vibeos kernel
evaluated here (15k LoC, 452 functions). The key architectural difference
is the verification oracle: \fmagent uses LLM reasoning (not sound for
arithmetic and memory bugs); \tool uses CBMC (sound within the unwinding
bound, producing concrete counterexamples). On \vibeos, \fmagent reports 23
bugs confirmed by live test-case execution; \tool reports 173 findings of
which 32 are runtime-confirmed (\dynconf). The counts are not directly
comparable: the tools target different bug classes (semantic/logic bugs
vs.\ low-level memory-safety and termination violations) and use different
confirmation mechanisms. They are best understood as complementary.
AutoSpec~\cite{autospec} synthesizes ACSL specifications for Frama-C/Wp by
coupling LLM generation with static analysis and a deductive checker in a
refinement loop. Lemur~\cite{lemur} integrates LLMs in automated program
verification, using LLMs to generate verification artifacts that a solver then
checks. Loop-invariant generation work~\cite{wu-invariants} similarly couples
LLM generation with formal solvers at the level of individual invariants.
\tool contributes to this line by targeting system-level verification with BMC
as the solver---enabling concrete counterexamples and direct loop
handling---using top-down call-graph traversal to derive caller-driven
specifications, and adding three-way counterexample classification with
compositional precondition refinement.

\noindent\textbf{Traditional driver and kernel verification.}
SLAM/SDV~\cite{slam,sdv} and BLAST check a fixed catalog of API-usage
properties over device drivers. Smatch~\cite{smatch} applies pattern-based
static analysis to the Linux kernel. These tools check predefined properties;
\tool's LLM-generated specs can in principle express semantic invariants that
pattern-based analyzers cannot, at the cost of depending on LLM judgment.

\noindent\textbf{Symbolic execution.}
KLEE~\cite{klee} achieves high path coverage via symbolic execution but
struggles with unbounded loops and large call graphs. Serval~\cite{serval}
applies symbolic evaluation to verified system kernels within decidable
fragments. Both differ from \tool in operating at the whole-function or
whole-system level rather than per-function with caller-derived specs.

\noindent\textbf{Deductive verification.}
Frama-C/Wp~\cite{framac} and related tools support full functional
correctness within deductive frameworks but require significant manual
annotation effort. \tool targets the annotation-free, automated end of the
spectrum, sacrificing completeness for scalability and concrete
counterexamples.

\noindent\textbf{CEGAR.}
Counterexample-guided abstraction refinement~\cite{cegar} is a classical
technique for iteratively refining abstract models when spurious
counterexamples are found. \tool applies the CEGAR idea at the
\emph{specification} level---refining LLM-generated specs rather than
predicate abstractions---and extends it with compositional propagation to
callers.

%% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
%% ─────────────────────────────────────────────────────────────────────────────

We have presented agentic model checking and \tool, a prototype
implementation that pairs LLM agents for specification generation,
counterexample classification, and precondition refinement with CBMC as a
sound verification backend. The central architectural principle---\emph{agents
propose, conventional tools dispose}---assigns tasks to LLMs where semantic
reasoning is appropriate while reserving all soundness-relevant decisions for
deterministic solvers.

Evaluated on \vibeos, \tool reports 173 confirmed findings across 37 kernel
modules, including 32 runtime-confirmed faults (SIGSEGV in compiled harnesses)
and one finding that independently cross-validates an open issue in the
\vibeos tracker. The FilteringOnly ablation shows that on this corpus,
refinement acts as a precision filter (eliminating 18 findings vs.\ 59 for
FilteringOnly) rather than a recall enhancer; resolving whether those
eliminated findings represent genuine precision improvement or over-refinement
artifacts is the most important next step.

\smallskip\noindent\textbf{Future work.}
The four most important open items are:
(i) a manual precision audit of a sampled subset of the 173 findings to
establish the classifier's false-positive rate;
(ii) running the CBMCAlone and AMCAblation baselines on \vibeos to complete
the quantitative comparison;
(iii) evaluating \tool on a broader corpus of LLM-generated drivers and
historical buggy kernel versions; and
(iv) activating the spec-quality defenses at scale to bound the over-weak-spec
failure mode.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
